NOPT048 Notes

### 2019-06-16 misc/study-notes/ndmi011notes.tex
Fixed many typos in graph theory notes

\proof The ``if'' direction is left as an exercise for the student. To prove the ``only if'' direction, we use the max-flow min-cut theorem. We construct a directed graph $G'$ from $G$:
$$\eqalign{
    V(G') &= V(G)\cr
    E(G') &= \{uv, vu : uv\in E(G)\}
}$$
To prove the vertex version of the theorem, for every vertex $v\in V'\setminus\{s,t\}$, we set a vertex capacity $d(v) = 1$. To prove the edge version of the theorem, we set $c(e) = 1$ for all edges in $E(G')$. Now we consider the vertex and edge cases separately.
\medskip
\item {1.} Assume that $st\notin E(G')$. Any s-t-vertex cut of $G'$ separates $G$; since any s-t-path in $G'$ is also a t-s-path, the s-t-vertex cut also separates $t$ from $s$. Let $T$ denote a minimum-cardinality s-t-vertex-cut in $G'$ and we get that $d(T) \geq \kappa(G)$. There exists a 0-1 flow with $|f|\geq \kappa(G)$. So at each vertex, the number of incoming edges is at most 1 and the number of outgoing edges is at most 1 as well. This implies that there are $|f|$ independent pats from $s$ to $t$. But by the max-flow min-cut theorem, $|f| = d(T) \geq \kappa(G) = k$. If it so happens that $st\in E(G)$, then $(s, st, t)$ is an additional independent s-t-path to those in $G-st$ and we know that $\kappa(G-st)\geq \kappa(G) - 1$. Hence $G$ being $k$-connected implies that there are at least $k$ independent paths between any $s, t$.
\smallskip
\item {2.} In the case of edge capacities, any s-t-edge-cut separates $G$ as any t-s-path is also an s-t-path in G'. Let $S$ denote a minimum s-t-edge cut. Then $c(E)\geq \lambda(G)$. Again, there exists a 0-1 flow f with $|f|\geq \lambda(G)$, but this time it may be possible that a vertex has many incoming and outgoing edges. However, we know that the number of incoming edges that carry flow must equal the number of outgoing edges that carry flow at any vertex. Let $H=(V,F)$ be a subgraph of $G$, where $F = \{e\in E(G) : f(e) = 1\}$. While there exists an s-t-path $P$ in $H$, we can remove any all edges of $P$ from $H$ while maintaining flow conservation at every vertex $v\in V\setminus\{s,t\}$. We can repeat the algorithm as many times as there are edges coming out of $S$, so at least $|f|$ times. Hence $|f|\geq \lambda(G)\geq k$.\slug

@@

\proclaim Lemma. Let $(X,{\cal L})$ be a finite projective plane. Then for all $L, L'\in {\cal L}$, there exists an $z\in L$ such that $x\notin L\cup L'$

@@

\proclaim Theorem. Let $(X,{\cal L})$ be a finite projective plane of order $n$. Then the following all hold:
\medskip
\item {1.} Exactly $n+1$ lines pass through any point $x\in X$.
\smallskip
\item {2.} $|X| = n^2 + n + 1$.
\smallskip
\item {3.} $|{\cal L}| = n^2 + n + 1$.
\medskip

@@

Let $x\in X$ be given and let $L$ be a line in ${\cal L}$ such that $x\notin L$. Now we may prove each part of the theorem.
\medskip
\item {1.} For all $y\in L$, there exists a line $\overline{xy}$ passing through $x$. There are $n+1$ points $y\in L$, so there are at least $n+1$ lines passing through $x$. Now for all $L'\in{\cal L}$ such that $x\in L'$, $|L\cap L'| = 1$ by (P1), so $L'$ was already counted above. So there are exactly $n+1$ lines through $x$.
\smallskip
\item {2.} Let $L_1 = \overline{xx_i}$ where $x_i\in L, i\in\{1,\ldots, n+1\}$. By (P1), $L_i\cap L_j = \{x\}$ for all $i\neq j$. As $|L_i\setminus\{x\}| = n$ we have
    $$|X|\geq |\bigcup_{i=1}^{n+1} L_i| = (n+1)n + 1 = n^2 + n + 1,$$
    so $|X|\geq n^2 + n + 1$.
    Now we show that for all $p\in X$, there exists an $i\in \{1,\ldots,n+1\}$ such that $p\in L_i$. If $p=x$. this is clearly true, so assume $p\neq x$. Then $\overline{px}\cap L = \{x_i\}$ for some $i$ by (P1). So $\overline{px} = L_i$ for some $i$ and we already counted $p$ above. So $|X| = n^2 + n + 1$.
\smallskip
\item {3.} This follows from item 2 and duality, which we will introduce next.\slug

The {\it incidence graph} of a finite projective plane $(X,{\cal L})$, is a bipartite graph $G = (V, E)$ with $V = X\cup {\cal L}$ and $E = \{xL : if x\in L\}$. The concept of duality involves switching the roles of $X$ and ${\cal L}$ in this bipartite graph, interpreting $X$ as lines and ${\cal L}$ as points.

@@

\proof We show that $({\cal L}, \Lambda)$ satisfies each of (P0), (P1), and (P2).
\medskip
\item {(P1)} We need to find lines $L_1, L_2, L_3, L_4$ such that any for any $\lambda\in \Lambda$, with $\lambda = \{L\in {\cal L} : x\in L\}$ for some $x\in X$, the point $x$ is contained in at most two of $L_1, L_2, L_3, L_4$. Let $F = \{a,b,c,d\}$ be the set given by applying (P0) to $(X,{\cal L})$. Let $L_1 = \overline{ab}, L_2 = \overline{cd}, L_3 = \overline{ad}, L_4 = \overline{bc}$. Suppose, towards a contradiction, that $x$ is in three of these lines. Without loss of generality, suppose $x\in L_1\cap L_2\cap L_3$. By (P1) for $(X,{\cal L})$, $|L_i\cap L_j|$ for all $i\neq j$. Since $x\in L_1\cap L_3$, we have that $x = a$. But since $x\in L_2\cap L_3$, we have that $x = d$. This is a contradiction, since $a\neq d$. The same is true for any other triple of these four lines.
\smallskip
\item {(P2)} Let $\lambda_1,\lambda_2\in \Lambda, \lambda_1\neq\lambda_2$ be given, i.e.\  $\lambda_1 = \{ L\in {\cal L} : x_1\in L\}$ and $\lambda_2 = \{L\in {\cal L} : x_2\in L\}$ for some points $x_1\neq x_2$. We need that $|\lambda_1\cap\lambda_2| = 1$, but this follows from (P2) for $(X,{\cal L})$, since $x_1$ and $x_2$ intersect at exactly one line.
\smallskip
\item {(P3)} Let $L_1, L_2\in {\cal L}$ with $L_1\neq L_2$. It suffices to show that there exists a unique $\lambda \in \Lambda$ such that $L_1\in \lambda$ and $L_2\in \lambda$. But since $\lambda = \{L\in {\cal L} : x\in L\}$ for some $x$, this is exactly (P1) for $(X,{\cal L})$.\slug

@@

So let $M = \{A_1, \ldots, A_t\}$ and for each $A_i\in M$, permute the entries such that the first row of the resulting Latin square $A'_i$ is $(1,2,3,\ldots,n)$. By our earlier observation, $A'_1,\ldots,A'_k$ are pairwise orthogonal. Now we zoom in on the second row, first column if $A'_i$ for some $i$. Call this cell $k$. First we notice that $k\neq 1$, since the first element of the first row is 1. Then since comparing any two $A'_i$ pairwise, we get all the pairs $(1,1), (2,2), \ldots, (n,n)$ in the first row, each of $2,3,\ldots,n$ can appear in the second row, first column of at most one $A'_i$. So $t = |M|\leq n-1$.\slug

@@

\proof By induction on $k+l$. In the base case $k = 1$ or $l=1$, all the theorem says is that there is an independent set or clique of size 4. For the inductive step, let $k,l\geq 2$ and assume that the claim holds for $k,l-1$ or $k-1,l$. By Pascal's formula,
$${k+l-2\choose k-1} = {k+l-3\choose k-1} + {k+l-3\choose k-2}.$$
Call the left-hand side $n$ and from the right-hand side, call the first summand $n_1$ and the second summand $n_2$. Now pick any vertex $u\in V$. Let $B = N(u)$ and $A = V\setminus(B\cup \{u\})$. First note that it is impossible that both $|A| < n_1$ and $|B|<n_2$ since that would imply that
$$n = 1 + |A| + |B| \leq 1 + (n_1-1) + (n_2=1) = n-1,$$
a contradiction. So either $|A|\geq n_1$ or $|B|\geq n_2$.

@@

Now we create a random graph $G$ by starting with a set of $n$ vertices and adding each edge with probablity $1/2$ uniformly at random. We're trying to prove that if $n\leq 2^{k/2}$ there is a $G$ with $|V| = n$ such that $\alpha(G) < k$ and $\omega(G) < k$, so it suffices to show that
$$\Pr[\alpha(G) \geq k\ \hbox{or}\ \omega(G)\geq k : n\leq k^{k/2}] < 1.$$
Let $A$ be a subset of vertices with $|A|=k$. Let $K_A$ denote the proposition ``A forms a clique'' and let $I_A$ denote the proposition ``A forms an independent set". Then let $X_A = K_A \cup I_A$. $\Pr[K_A] = (1/2)^{k\choose 2} = \Pr[I_A]$ and since $K_A \cap I_A = \emptyset$,
$$X_A = 2\cdot\left({1\over 2}\right)^{k\choose 2} = 2^{1-{k\choose 2}}.$$
Let $\Pr(Y)$ denote the probability that $\alpha(G) \geq k$ or $\omega(G) \geq k$. This is equal to the probablilty that $X_A$ holds for some $A\subseteq V$ with $|A|\geq k$. So
$$\Pr[Y] \leq \sum_{A\subseteq V, |A|\geq k} 2^{1-{k\choose 2}} = {n\choose 2}2^{1-{k\choose 2}} \leq {n^k\over k!}2^{1-{k\choose 2}},$$
and we can further derive the strict inequality
$${n^k\over k!}2^{1-{k\choose 2}} < {n^k\over {2^{k/2+1}}}2^{1-(k^2/2-k/2)} = \left({n\over 2^{k/2}}\right)^k,$$
which is less than $1$, since $n\leq 2^{k/2}$. \slug

@@

\proof Suppose, towards a contradiction, that every increasing/decreasing subsequence has length at most $r-1$/$s-1$ respectively. Then we label each $x_i$ of $S$ with a pair $(a_i, b_i)$ where $a_i$/$b_i$ is the length of the longest increasing/decreasing subsequence ending in $x_i$. (Note that each of the numbers $x_i, x_j$ have different labels.) Without loss of generality, assume $i<j$. If $x_i\leq x_j$, then $a_i < a_j$, as any increasing subsequence ending in $x_i$ can be extended by $x_j$. Likewise, if $x_j\leq x_i$, then $a_j < a_i$, as any increasing subsequence ending in $x_j$ can be extended by $x_i$.

@@

Suppose we have two people who need to communicate over a channel that is ``noisy'' in the sense that errors may be introduced into the message. To detect and solve this issue, checksums may be used. Suppose we have an alphabet $\Sigma = \{0,\ldots,q-1\}$ where $q$ is prime. Let $x$ denote a message $x\in \Sigma^k$ for some positive integer $k$. Then the {\it checksum} $c$ of $x$ is $\sum_{i=1}^k x_i \pmod{q}$. Then the message becomes $y\in {\bf F}_q^{k+1}$ where
$$ y_i = \cases{
    x_i & for $i\leq k$ \cr
    c   & for $i = k+1$
}$$
Since $q$ is prime, for any $x,y\in {\bf F}_q^kk$, if $x$ and $y$ differ at {\it exactly one} place, then their checksums will differ. But what happens if more errors occur? Also, is it possible for the receiver to determine where the error occured without asking the sender to resend the message?

@@

A {\it linear} code $C\subseteq {\bf F}_q^n$ is a linear subspace of ${\bf F}_q$, i.e.\  for all $x,y\in C$, $\alpha\in {\bf F}_q$, $\alpha x+y\in C$. There exist $k$ independent vectors $x_1,\ldots, x_k\in {\bf F}_q^n$ such that
$$C = \{\sum_{i=1}^k \alpha_ix_i : \alpha_i,\ldots \alpha_n\in {\bf F}_q\}.$$
More succinctly, we can express $C$ using the generator matrix $G\in {\bf F}_q^{n\times k}$. The matrix $G = \matrix{ x_1 & \cdots & x_k }$ and $C = \{G_\alpha : \alpha \in {\bf F}_q^k\}$. (Multiplying any message by $G$ on the left encodes it.)

Alternatively, represent $C$ by the {\it null space}. Recall that for any $k$-dimensional linear subspace $C$, there is an $(n-k)$-dimensional linear subspace $C^\perp$ such that for all $x\in C, y\in C^\perp$, $x^Ty = 0$. Using the generator matrix $H\in {\bf F}_k^{n\times (n-k)}$ of $C^\perp$, express $C = \{ x : x^TH = 0\}$. Then $H$ is called the {\it parity check} matrix of $C$. In a sense, linear codes are simple because we can detect errors simply by computing $x^TH$.

The {\it support} of a vector $x$ is the set $\{i : x_i \neq 0\}$ and the {\it Hamming weight} of $x$ is $\hbox{wt} x = |\{i : x_i \neq 0\}|$.

\proclaim Theorem. For a linear code $C$,
$$ \Delta(C) = \min\{\hbox{wt} x : x\in C\ \hbox{and}\ x\neq 0\}.$$

\proof For ease of notation, let $d$ denote $\min\{\hbox{wt} x : x\in C\ \hbox{and}\ x\neq 0\}$. First we show that $d\leq \Delta(C)$. Let $y,z\in C$ be such that $\Delta(y,z)$ = $\Delta(C)$. Since $C$ is linear, $x = y-z \in C$. Note that $\Delta(y,z) = |\{i : x_i \neq 0\}|$, so $d \leq \hbox{wt} x = \Delta(C)$.

Now we show that $d\geq \Delta(C)$. Note that $0\in C$. Let $x\neq 0$ be such that $\hbox{wt} x = d$. Then $\Delta(C)\leq \Delta(0,x) = |\{i : x_i \neq 0\}| = \hbox{wt} x = d$.\slug

We can use this theorem to construct a linear code with $\Delta(C) \geq 3$. We find a parity check matrix $H$ such that $x^TH \neq 0$ for any non-zero $x$ with $\hbox{wt} x\in \{1,2\}$. For simplicity, we will work with $q=2$. If $\hbox{wt} x = 1$, then $X$ has exactly one 1-entry, say it is at position $i$. So $x^TH$ is the $i$-th row $H_i$ of $H$. We need that $H_i$ is non-zero for every $i$. If $\hbox{wt} x = 2$, then $x^T$ has exactly two non-zero elements, say they are at positions $i$ and $j$. Then $x^TH$ = $(H_i + H_j)^T$ hence we need $H_i\neq H_j$ for all $i\neq j$.

@@

\thmparens{S}{Hamming bound} If an $(n,k,d)_q$-code exists, then
$$q^k\cdot\hbox{vol}\left(\lfloor {d-1\over 2} \rfloor, n\right) \leq q^n.$$

\proof Let $r = \lfloor (d-1)/2 \rfloor$ and consider $\bigcup_{x\in C} B(x,r)\subset \{1,\ldots,q\}^n$. By observation,
$$|\bigcup_{x\in C} B(x,r)| = \sum_{x\in C} \hbox{vol}(r,n) = q^k\cdot \hbox{vol}(r,n) \leq q_n.\noskipslug$$

Note that if $d = 3$ and $q = 2$ then the Hamming bound is $2^k(n+1)\leq 2^n$ and the Hamming code matches this bound with equality as $k=n-log_2(n+1)$. An $(n,k,d)_2$-code is called {\it perfect} if $q^k\cdot\hbox{vol}(\lfloor (d-1)/2 \rfloor, n) = q^n$. There are no codes with larger size $k$ than a $(n, n-d+1, d)_2$-code (even if we let $q>2$), as the following theorem shows.

>>>>

\proof The ``if'' direction is left as an exercise for the student. To prove the ``only if'' direction, we use the max-flow min-cut theorem. We construct a directed graph $G'$ from $G$:
$$\eqalign{
    V(G') &= V(G)\cr
    E(G') &= \{uv, vu : uv\in E(G)\}
}$$
To prove the vertex version of the theorem, for every vertex $v\in V'\setminus\{s,t\}$, we set a vertex capacity $d(v) = 1$. To prove the edge version of the theorem, we set $c(e) = 1$ for all edges in $E(G')$. Now we consider the vertex and edge cases separately.
\medskip
\item {1.} Assume that $st\notin E(G')$. Any s-t-vertex cut of $G'$ separates $G$; since any s-t-path in $G'$ is also a t-s-path, the s-t-vertex cut also separates $t$ from $s$. Let $T$ denote a minimum-cardinality s-t-vertex-cut in $G'$ and we get that $d(T) \geq \kappa(G)$. There exists a 0-1 flow with $|f|\geq \kappa(G)$. So at each vertex, the number of incoming edges is at most 1 and the number of outgoing edges is at most 1 as well. This implies that there are $|f|$ independent paths from $s$ to $t$. But by the max-flow min-cut theorem, $|f| = d(T) \geq \kappa(G) = k$. If it so happens that $st\in E(G)$, then $(s, st, t)$ is an additional independent s-t-path to those in $G-st$ and we know that $\kappa(G-st)\geq \kappa(G) - 1$. Hence $G$ being $k$-connected implies that there are at least $k$ independent paths between any $s, t$.
\smallskip
\item {2.} In the case of edge capacities, any s-t-edge-cut separates $G$ as any t-s-path is also an s-t-path in G'. Let $S$ denote a minimum s-t-edge cut. Then $c(E)\geq \lambda(G)$. Again, there exists a 0-1 flow f with $|f|\geq \lambda(G)$, but this time it may be possible that a vertex has many incoming and outgoing edges. However, we know that the number of incoming edges that carry flow must equal the number of outgoing edges that carry flow at any vertex. Let $H=(V,F)$ be a subgraph of $G$, where $F = \{e\in E(G) : f(e) = 1\}$. While there exists an s-t-path $P$ in $H$, we can remove any all edges of $P$ from $H$ while maintaining flow conservation at every vertex $v\in V\setminus\{s,t\}$. We can repeat the algorithm as many times as there are edges coming out of $S$, so at least $|f|$ times. Hence $|f|\geq \lambda(G)\geq k$.\slug

@@

\proclaim Lemma. Let $(X,{\cal L})$ be a finite projective plane. Then for all $L, L'\in {\cal L}$, there exists an $z\in L$ such that $z\notin L\cup L'$

@@

\proclaim Theorem. Let $(X,{\cal L})$ be a finite projective plane of order $n$. Then the following statements all hold:
\medskip
\item {1.} Exactly $n+1$ lines pass through any point $x\in X$.
\smallskip
\item {2.} $|X| = n^2 + n + 1$.
\smallskip
\item {3.} $|{\cal L}| = n^2 + n + 1$.
\medskip

@@

Let $x\in X$ be given and let $L$ be a line in ${\cal L}$ such that $x\notin L$. Now we may prove each part of the theorem.
\medskip
\item {1.} For all $y\in L$, there exists a line $\overline{xy}$ passing through $x$. There are $n+1$ points $y\in L$, so there are at least $n+1$ lines passing through $x$. Now for all $L'\in{\cal L}$ such that $x\in L'$, $|L\cap L'| = 1$ by (P1), so $L'$ was already counted above. So there are exactly $n+1$ lines through $x$.
\smallskip
\item {2.} Let $L_i = \overline{xx_i}$ where $x_i\in L, i\in\{1,\ldots, n+1\}$. By (P1), $L_i\cap L_j = \{x\}$ for all $i\neq j$. As $|L_i\setminus\{x\}| = n$ we have
    $$|X|\geq |\bigcup_{i=1}^{n+1} L_i| = (n+1)n + 1 = n^2 + n + 1,$$
    so $|X|\geq n^2 + n + 1$.
    Now we show that for all $p\in X$, there exists an $i\in \{1,\ldots,n+1\}$ such that $p\in L_i$. If $p=x$, this is clearly true, so assume $p\neq x$. Then $\overline{px}\cap L = \{x_i\}$ for some $i$ by (P1). So $\overline{px} = L_i$ for some $i$ and we already counted $p$ above. So $|X| = n^2 + n + 1$.
\smallskip
\item {3.} This follows from item 2 and duality, which we will introduce next.\slug

The {\it incidence graph} of a finite projective plane $(X,{\cal L})$, is a bipartite graph $G = (V, E)$ with $V = X\cup {\cal L}$ and $E = \{xL : \hbox{for all pairs}\ x,L\ \hbox{such that}\ x\in L\}$. The concept of duality involves switching the roles of $X$ and ${\cal L}$ in this bipartite graph, interpreting $X$ as lines and ${\cal L}$ as points.

@@

\proof We show that $({\cal L}, \Lambda)$ satisfies each of (P0), (P1), and (P2).
\medskip
\item {P1)} We need to find lines $L_1, L_2, L_3, L_4$ such that any for any $\lambda\in \Lambda$, with $\lambda = \{L\in {\cal L} : x\in L\}$ for some $x\in X$, the point $x$ is contained in at most two of $L_1, L_2, L_3, L_4$. Let $F = \{a,b,c,d\}$ be the set given by applying (P0) to $(X,{\cal L})$. Let $L_1 = \overline{ab}, L_2 = \overline{cd}, L_3 = \overline{ad}, L_4 = \overline{bc}$. Suppose, towards a contradiction, that $x$ is in three of these lines. Without loss of generality, suppose $x\in L_1\cap L_2\cap L_3$. By (P1) for $(X,{\cal L})$, $|L_i\cap L_j|$ for all $i\neq j$. Since $x\in L_1\cap L_3$, we have that $x = a$. But since $x\in L_2\cap L_3$, we have that $x = d$. This is a contradiction, since $a\neq d$. The same is true for any other triple of these four lines.
\smallskip
\item {P2)} Let $\lambda_1,\lambda_2\in \Lambda, \lambda_1\neq\lambda_2$ be given, i.e.\  $\lambda_1 = \{ L\in {\cal L} : x_1\in L\}$ and $\lambda_2 = \{L\in {\cal L} : x_2\in L\}$ for some points $x_1\neq x_2$. We need that $|\lambda_1\cap\lambda_2| = 1$, but this follows from (P2) for $(X,{\cal L})$, since $x_1$ and $x_2$ intersect at exactly one line.
\smallskip
\item {P3)} Let $L_1, L_2\in {\cal L}$ with $L_1\neq L_2$. It suffices to show that there exists a unique $\lambda \in \Lambda$ such that $L_1\in \lambda$ and $L_2\in \lambda$. But since $\lambda = \{L\in {\cal L} : x\in L\}$ for some $x$, this is exactly (P1) for $(X,{\cal L})$.\slug

@@

So let $M = \{A_1, \ldots, A_t\}$ and for each $A_i\in M$, permute the $n$ columns such that the first row of the resulting Latin square $A'_i$ is $(1,2,3,\ldots,n)$. By our earlier observation, $A'_1,\ldots,A'_k$ are pairwise orthogonal. Now we zoom in on the second row, first column if $A'_i$ for some $i$. Call this cell $k$. First we notice that $k\neq 1$, since the first element of the first row is 1. Then since comparing any two $A'_i$ pairwise, we get all the pairs $(1,1), (2,2), \ldots, (n,n)$ in the first row, each of $2,3,\ldots,n$ can appear in the second row, first column of at most one $A'_i$. So $t = |M|\leq n-1$.\slug

@@

\proof By induction on $k+l$. In the base case $k = 1$ or $l=1$, all the theorem says is that there is an independent set or clique of size 4. For the inductive step, let $k,l\geq 2$ and assume that the claim holds for $k,l-1$ or $k-1,l$. By Pascal's formula,
$${k+l-2\choose k-1} = {k+l-3\choose k-1} + {k+l-3\choose k-2}.$$
Call the left-hand side $n$ and from the right-hand side, call the first summand $n_1$ and the second summand $n_2$. Now pick any vertex $u\in V$. Let $B = N(u)$ and $A = V\setminus(B\cup \{u\})$. First note that it is impossible that both $|A| < n_1$ and $|B|<n_2$ since that would imply that
$$n = 1 + |A| + |B| \leq 1 + (n_1-1) + (n_2-1) = n-1,$$
a contradiction. So either $|A|\geq n_1$ or $|B|\geq n_2$.

@@

Now we create a random graph $G$ by starting with a set of $n$ vertices and adding each edge with probablity $1/2$ uniformly at random. We're trying to prove that if $n\leq 2^{k/2}$ there is a $G$ with $|V| = n$ such that $\alpha(G) < k$ and $\omega(G) < k$, so it suffices to show that
$$\Pr[\alpha(G) \geq k\ \hbox{or}\ \omega(G)\geq k : n\leq k^{k/2}] < 1.$$
Let $A$ be a subset of vertices with $|A|=k$. Let $K_A$ denote the proposition ``A forms a clique'' and let $I_A$ denote the proposition ``A forms an independent set". Then let $X_A = K_A \cup I_A$. $\Pr[K_A] = (1/2)^{k\choose 2} = \Pr[I_A]$ and since $K_A \cap I_A = \emptyset$,
$$X_A = 2\cdot\left({1\over 2}\right)^{k\choose 2} = 2^{1-{k\choose 2}}.$$
Let $\Pr(Y)$ denote the probability that $\alpha(G) \geq k$ or $\omega(G) \geq k$. This is equal to the probablilty that $X_A$ holds for some $A\subseteq V$ with $|A|\geq k$. So
$$\Pr[Y] \leq \sum_{A\subseteq V\atop |A|\geq k} 2^{1-{k\choose 2}} = {n\choose 2}2^{1-{k\choose 2}} \leq {n^k\over k!}2^{1-{k\choose 2}},$$
and we can further derive the strict inequality
$${n^k\over k!}2^{1-{k\choose 2}} < {n^k\over {2^{k/2+1}}}2^{1-(k^2/2-k/2)} = \left({n\over 2^{k/2}}\right)^k,$$
which is less than $1$, since $n\leq 2^{k/2}$. \slug

@@

\proof Suppose, towards a contradiction, that every increasing/decreasing subsequence has length at most $r-1$/$s-1$ respectively. Then we label each $x_i$ of $S$ with a pair $(a_i, b_i)$ where $a_i$/$b_i$ is the length of the longest increasing/decreasing subsequence ending in $x_i$.

We claim that each of the numbers $x_i, x_j$ have different labels. Let $i\neq j$ and without loss of generality, assume $i<j$. If $x_i\leq x_j$, then $a_i < a_j$, as any increasing subsequence ending in $x_i$ can be extended by $x_j$. Likewise, if $x_i\geq x_j$, then $b_j > b_i$, as any decreasing subsequence ending in $x_i$ can be extended by $x_j$.

@@

Suppose we have two people who need to communicate over a channel that is ``noisy'' in the sense that errors may be introduced into the message. To detect and solve this issue, checksums may be used. Suppose we have an alphabet $\Sigma = \{0,\ldots,q-1\}$ where $q$ is prime. Let $x$ denote a message $x\in \Sigma^k$ for some positive integer $k$. The {\it checksum} $c$ of $x$ is $\sum_{i=1}^k x_i \pmod{q}$ and we can encode te message as $y\in {\bf F}_q^{k+1}$ where
$$ y_i = \cases{
    x_i & for $i\leq k$ \cr
    c   & for $i = k+1$
}$$
Since $q$ is prime, for any $x,y\in {\bf F}_q^kk$, if $x$ and $y$ differ at {\it exactly one} place, then their checksums will differ. But what happens if more errors occur? Also, is it possible for the receiver to determine where the error occured without asking the sender to resend the message?

@@

\smallskip

A {\it linear} code $C\subseteq {\bf F}_q^n$ is a linear subspace of ${\bf F}_q$, i.e.\  for all $x,y\in C$, $\alpha\in {\bf F}_q$, $\alpha x+y\in C$. There exist $k$ independent vectors $x_1,\ldots, x_k\in {\bf F}_q^n$ such that
$$C = \bigg\{\sum_{i=1}^k \alpha_ix_i : \alpha_i,\ldots \alpha_n\in {\bf F}_q\bigg\}.$$
More succinctly, we can express $C$ using the generator matrix $G\in {\bf F}_q^{n\times k}$. The matrix $G = \matrix{ x_1 & \cdots & x_k }$ and $C = \{G_\alpha : \alpha \in {\bf F}_q^k\}$. (Multiplying any message by $G$ on the left encodes it.)

Alternatively, we may represent $C$ as the {\it null space} of a matrix. Recall that for any $k$-dimensional linear subspace $C$, there is an $(n-k)$-dimensional linear subspace $C^\perp$ such that for all $x\in C, y\in C^\perp$, $x^Ty = 0$. Using the generator matrix $H\in {\bf F}_k^{n\times (n-k)}$ of $C^\perp$, express $C = \{ x : x^TH = 0\}$. Then $H$ is called the {\it parity check} matrix of $C$. In a sense, linear codes are simple because we can detect errors simply by computing $x^TH$.

The {\it support} of a vector $x$ is the set $\{i : x_i \neq 0\}$ and the {\it Hamming weight} of $x$ is $\hbox{wt}\,x = |\{i : x_i \neq 0\}|$.

\proclaim Theorem. For a linear code $C$,
$$ \Delta(C) = \min\{\hbox{wt}\,x : x\in C\ \hbox{and}\ x\neq 0\}.$$

\proof For ease of notation, let $d$ denote $\min\{\hbox{wt}\,x : x\in C\ \hbox{and}\ x\neq 0\}$. First we show that $d\leq \Delta(C)$. Let $y,z\in C$ be such that $\Delta(y,z)$ = $\Delta(C)$. Since $C$ is linear, $x = y-z \in C$. Note that $\Delta(y,z) = |\{i : x_i \neq 0\}|$, so $d \leq \hbox{wt}\,x = \Delta(C)$.

Now we show that $d\geq \Delta(C)$. Note that $0\in C$. Let $x\neq 0$ be such that $\hbox{wt}\,x = d$. Then $\Delta(C)\leq \Delta(0,x) = |\{i : x_i \neq 0\}| = \hbox{wt}\,x = d$.\slug

We can use this theorem to construct a linear code with $\Delta(C) \geq 3$. We find a parity check matrix $H$ such that $x^TH \neq 0$ for any non-zero $x$ with $\hbox{wt} x\in \{1,2\}$. For simplicity, we will work with $q=2$. If $\hbox{wt}\,x = 1$, then $X$ has exactly one 1-entry, say it is at position $i$. So $x^TH$ is the $i$-th row $H_i$ of $H$. We need that $H_i$ is non-zero for every $i$. If $\hbox{wt}\,x = 2$, then $x^T$ has exactly two non-zero elements, say they are at positions $i$ and $j$. Then $x^TH$ = $(H_i + H_j)^T$ hence we need $H_i\neq H_j$ for all $i\neq j$.

@@

\thmparens{S}{Hamming bound} If an $(n,k,d)_q$-code exists, then
$$q^k\cdot\hbox{\rm vol}\left(\left\lfloor {d-1\over 2} \right\rfloor, n\right) \leq q^n.$$

\proof Let $r = \lfloor (d-1)/2 \rfloor$ and consider $\bigcup_{x\in C} B(x,r)\subset \{1,\ldots,q\}^n$. By observation,
$$\left|\bigcup_{x\in C} B(x,r)\right| = \sum_{x\in C} \hbox{vol}\,(r,n) = q^k\cdot \hbox{vol}\,(r,n) \leq q^n.\noskipslug$$

Note that if $d = 3$ and $q = 2$ then the Hamming bound is $2^k(n+1)\leq 2^n$ and the Hamming code matches this bound with equality as $k=n-log_2(n+1)$. An $(n,k,d)_2$-code is called {\it perfect} if $q^k\cdot\hbox{vol}\,(\lfloor (d-1)/2 \rfloor, n) = q^n$. There are no codes with larger size $k$ than a $(n, n-d+1, d)_2$-code (even if we let $q>2$), as the following theorem shows.
